Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Jupiter-Sized Planet in IC~2602

### 2020-06-22 paper/ms.tex
update manuscript

@@

\title{
  Cluster Difference Imaging Photometric Survey (CDIPS) II:
  Validatation of a Jupiter-Sized Planet in IC~2602
}

@@

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS), the ESA
  Gaia mission, ground-based photometry, and spectroscopy from
  CTIO1.5/CHIRON, MPG2.2/FEROS, and AAT/Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G-dwarf in IC~2602 (the ``Southern
  Pleiades'').  The star and validated planet are therefore 30$\pm$20
  million years old.  The planet is warm ($P = 8.YY\,{\rm d}$) and
  roughly Jupiter-sized.  However, its transits are grazing: $b >
  0.92$ at 3$\sigma$.  From TESS photometry alone, the planetary size
  lies within 0.58--3.87 $R_{\rm Jup}$, due to the degeneracy between
  the planetary size and the impact parameter of the transit.  From
  radial velocity monitoring, we limit the mass of planet to less than
  2.XX $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are cause for
  concern, as they usually are indicative of astrophysical false
  positive scenarios.  Our follow-up data render such scenarios highly
  unlikely.  Multi-color photometry requires eclipsing companions to
  have nearly identical color as TOI$\,$837, eliminating hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios
  are limited by speckle imaging, but formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent exoplanet
  orbiting a bright host star.
  Further analysis of its mass, atmosphere, and obliquity should
  confirm its planetary nature beyond doubt, and are worthwhile
  because young planets are expected to do interesting things.
\end{abstract}

%FIXME these keywords.
\keywords{
	Exoplanets (XXX),
	Exoplanet evolution (491),
	Stellar ages (1581),
	Young star clusters (XXX)
}

@@

Recently, young planets have been hype.

%TODO: pull from the cdips repo! I already wrote a lot of this...

TESS \citep{ricker_transiting_2015}.

Section~\ref{sec:observations} describes the identification of the
candidate, and the follow-up observations collected.
Section~\ref{sec:validation} combines these data to assess the 
system's false positive probability, and finds that \pn\ is almost
certainly a planet.  Section~\ref{sec:system} presents our knowledge
of the cluster (Section~\ref{subsec:cluster}), the star (Section~\ref{subsec:star})
and the planet (Section~\ref{subsec:planet}).  We conclude
with a discussion in
Section~\ref{sec:discussion}.

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf TESS lightcurve of \tn\ (Sectors 10 and 11).}
    {\it Top}: \texttt{PDCSAP} mean-subtracted relative flux at
    2-minute sampling. Spot-induced stellar variability is the
    dominant signal.  Dashed lines show the five transits observed by
    TESS.
    {\it Bottom}: Zoomed windows of individual transits.  Red points
    show manually identified stellar flares.
		\label{fig:rawzoom}
	}
\end{figure*}

%TODO: the scale of this plot is off. (Different font-sizes from
%subplots).
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f2a.pdf}
		}

>>>>

\newcommand{\stscilink}{\textsc{\url{archive.stsci.edu/hlsp/cdips}}}
\newcommand{\datasetlink}{\textsc{\dataset[doi.org/10.17909/t9-ayd0-k727]{https://doi.org/10.17909/t9-ayd0-k727}}}

@@

\title{
  Cluster Difference Imaging Photometric Survey (CDIPS) II.
  TOI$\,$837: A Validated Jupiter-Sized Planet in IC~2602
}

@@

% TODO: SpType subtype?
\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS), the ESA
  Gaia mission, ground-based photometry, and spectroscopy from
  CTIO1.5/CHIRON, MPG2.2/FEROS, and AAT/Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G-dwarf in IC~2602 (the ``Southern
  Pleiades'').  The star and validated planet are therefore 30$\pm$20
  million years old.  The planet is warm ($P = 8.32\,{\rm d}$) and
  roughly Jupiter-sized.  However, its transits are grazing: $b >
  0.92$ at 3$\sigma$.  From TESS photometry alone, the planetary size
  lies within 0.58--3.87 $R_{\rm Jup}$, due to the degeneracy between
  the planetary size and the impact parameter of the transit.  From
  radial velocity monitoring, we limit the mass of planet to less than
  2.XX $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are cause for
  concern: they usually are indicative of astrophysical false
  positive scenarios.  Our follow-up data show that such scenarios are
  highly unlikely.  Multi-color photometry requires eclipsing
  companions to have nearly identical color as TOI$\,$837, ruling out
  hierarchical eclipsing binary scenarios.  Background eclipsing
  binary scenarios are limited by speckle imaging, but formally remain
  a 0.2\% possibility.  TOI 837.01 is therefore a validated adolescent
  exoplanet.  Further observations of its stellar obliquity,
  atmosphere, and mass should confirm its planetary nature beyond
  doubt, and could improve our understanding of how the physical and
  orbital properties of exoplanets change in time.
\end{abstract}

\keywords{
	Exoplanets (498),
  Transits (1711),
	Exoplanet evolution (491),
	Stellar ages (1581),
	Young star clusters (1833)
}

@@

Over the first one hundred million years of their lives, exoplanet
systems are expected to undergo major physical and dynamical changes.
For a typical Sun-like star, the protoplanetary disk disperses within
$\sim$3 million years (CITE IR + sub-mm results).  The planet
formation process presumably finishes within this time for gas giants
(CITE).  The rocky planet population will also be mostly evolved
within $\sim$3 million years (CITE), but may undergo significant
changes over the next $\sim$10-100 million years through giant
impacts.  The Moon, for instance, is thought to have formed when the
solar system was XX million years old through such an impact (CITE).

A number of other processes are thought to happen relatively quickly.
Physical changes in the sizes of planets are expected as they
contract, and as their atmospheres undergo photoevaporation
\citep[{\it
e.g.},][]{Fortney_et_al_2007,Owen_Wu_2013,Fulton_et_al_2017}.  The
relative importance of these two processes is set by the planetary
mass, as well as the radiation environment.  Photoevaporation in
particular can be directly observed in {\it e.g.}, the He 10830\AA\
line (CITE, CITE, He-I or He-II?).  Dynamical changes are expected in the semi-major
axes, eccentricities, and stellar obliquities of young planets.  When
the gas disk is present, the planetary semi-major axis is thought to
change due to viscous coupling (CITE).  High-eccentricity migration
processes including planet-planet scattering (CITE), secular chaos
(CITE), and Kozai-Lidov oscillations (CITE) can also occur.  The
circularization timescale is thought to be such that for any giant
planets that migrated early, their orbits should circularize within
100 million years (CITE).

Finding and studying systems undergoing these evolutionary changes is
a major goal in contemporary exoplanet research.  To identify stars
younger than say 1$\,$Gyr, a number of direct and indirect methods
are available \citep{soderblom_ages_2010}.  The traditional approach
is to isochronally age-date coeval groups of stars, hereafter ``clusters''
\citep[{\it
e.g.},][]{lada_embedded_2003,zuckerman_young_2004,krumholz_star_2019}.
Young field stars can also be identified using age indicators such as
stellar rotation periods, the abundance of photospheric lithium, and
chromospheric diagnostics such as calcium emission and broadband UV
emission.  Studies by for instance \citet{david_discovery_2018} and
G{.}~Zhou et al. (2020, submitted) have combined these methods for
individual field stars.  Many of the latter methods were summarized by
\citet{mamajek_improved_2008}, and have since been calibrated by
{\it e.g.},
\citet{barnes_color-period_2015,meibom_spin-down_2015,angus_calibrating_2015}
and \citet{curtis_tess_2019} for stellar rotation,
\citet{zerjal_chromospherically_2017} for chromospheric activity, and
{\it e.g.}, \citet{berger_identifying_2018} and
\citet{zerjal_galah_2019} for lithium abundances.

A few dozen planets in clusters have been detected, and fewer
still have been closely characterized.  Despite the challenges of starspot-induced radial
velocity (RV) variations, RV surveys found success in the
Hyades, NGC~2523, Praesepe, and M~67
\citep{Sato_et_al_2007,lovis_mayor_2007,Quinn_et_al_2012,Malavolta_et_al_2016,brucalassi_search_2017}.
RV surveys of highly active pre-main sequence stars in Taurus also led to the
youngest hot Jupiters yet reported
\citep{donati_hj_2016,johns-krull_candidate_2016,biddle_k2_2018,flagg_co_2019}.

The transit method was comparatively slow to catch up.  Early deep
transit searches of open clusters by many groups did not yield
definitive planet detections
\citep{mochejska_planets_2005,mochejska_planets_2006,burke_survey_2006,aigrain_monitor_2007,irwin_monitordata_2007,miller_monitor_2008,pepper_photometric_2008,hartman_MMT_IV_2009}.
These searches were typically sensitive to planets larger than
Jupiter, on $\lesssim 3$ day orbital periods.  Hot Jupiter occurrence
rate limits were derived at the $\lesssim 5\%$ level \citep[{\it
e.g.},][]{burke_survey_2006,hartman_MMT_IV_2009}.  The modern 0.5-1\%
occurrence rate suggests that these early transit surveys needed a
greater data volume, at higher precision for detection to be possible
\citep{mayor_harps_2011,wright_frequency_2012,howard_planet_2012,petigura_metallicity_2018}.

Kepler observed a large enough number of stars 
with sufficient baseline and precision to detect transiting planets in
open clusters:  Kepler-66b and 67b, in the gigayear-old NGC~6811
\citep{borucki_kepler_2010,Meibom_et_al_2013}.  Though a broken reaction wheel ended the
prime Kepler mission, the repurposed K2 \citep{howell_k2_2014}
switched between fields along the ecliptic every quarter-year, and was
able to observe far more clusters and young stars.

The discoveries made by K2 through its surveys of Taurus, the Hyades,
Praesepe, and Upper Sco were a major inspiration for the present work
\citep[{\it
e.g.},][]{Mann_K2_25_2016,obermeier_k2_2016,Mann_et_al_2017,vanderburg_zeitVII_2018,ciardi_k2-136_2018,livingston_three_2018,mann_ZEITVI_2018,rizzuto_zeitVIII_2018,livingston_k2-264_2019}.
Observations with K2 convincingly showed that at least some close-in
planets must form within about 10 Myr
\citep{Mann_K2_33b_2016,David_et_al_2017}.  They also led to the first
hints that young planets may in fact be qualtiatively different from
their field counterparts.  For instance, based on its observed mass,
radius, and UV environment, the 700$\,$Myr K2-100b is probably
actively losing its atmosphere, and should become a bare rocky planet
over the next few hundred Myr
\citep{Mann_et_al_2017,barragan_radial_2019}.  The four transiting
planets in V1298~Tau (23$\,$Myr) are also likely to be actively
photoevaporating, and could represent a precursor to Kepler's compact
multiple systems \citep{david_four_2019,david_warm_2019}.

With the aim of advancing the young planet census, we have been using
data from the TESS spacecraft \citep{ricker_transiting_2015} to
perform a Cluster Difference Imaging Photometric Survey
\citep[CDIPS;][]{bouma_cluster_2019}.  Our targets in this survey are
candidate young stars that have been reported in the literature.  At
the time writing, of order $6\times10^5$ light curves from the first
year of TESS observations have been reduced, and are available at
MAST\footnote{\stscilink} through \datasetlink.  Searching these light
curves yielded the candidate transiting planet, \pn, that is the
subject of this analysis.

Section~\ref{sec:observations} describes the identification of the
candidate, and the follow-up observations collected.
Section~\ref{sec:validation} combines these data to assess the
system's false positive probability, and finds that \pn\ is almost
certainly a planet.  Section~\ref{sec:system} presents our knowledge
of the cluster (Section~\ref{subsec:cluster}), the star
(Section~\ref{subsec:star}) and the planet
(Section~\ref{subsec:planet}).  We conclude by discussing avenues for
improved characterization in Section~\ref{sec:discussion}.

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf TESS lightcurve of \tn\ (Sectors 10 and 11).}
    {\it Top}: \texttt{PDCSAP} median-subtracted relative flux at
    2-minute sampling in units of parts-per-thousand ($\times 10^{-3}$).
    Spot-induced stellar variability is the
    dominant signal.  Dashed lines show the five transits observed by
    TESS.
    {\it Bottom}: Zoomed windows of individual transits.  Red points
    highlight manually identified stellar flares.
		\label{fig:rawzoom}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f2a.pdf}
		}
